### 2022-10-18 paper.tex

% Template for ICASSP-2021 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}
\input{preamble.tex}

% Title.
% ------
\title{SOMETHING ABOUT THE DISTRIBUTED AVERAGING BASED APPROXIMATION IN SENSOR NETWORKS}
%
% Single address.
% ---------------
\name{Matthias Blochberger\(^1\), Filip Elvander\(^2\), Toon, ...\thanks{This research work was carried out at the ESAT Laboratory of KU Leuven, in the frame of the SOUNDS European Training Network. This project has received funding from the European Union's Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement No.\,956369. This research received funding in part from the European Union's Horizon 2020 research and innovation program / ERC Consolidator Grant: SONORA (No.\,773268). This paper reflects only the authors' views and the Union is not liable for any use that may be made of the contained information.}}
\address{\(^1\)KU Leuven, Dept. of Electrical Engineering (ESAT), STADIUS, 3001 Leuven, Belgium\\\(^2\)Aalto University, Dept. of Signal Processing and Acoustics, 02150 Espoo, Finland}

\begin{document}
\ninept
%
\maketitle
%
\begin{abstract}
    Distributed signal-processing algorithms in (wireless) sensor networks often aim to decentralize processing tasks as to reduce communication cost and computational complexity or avoid reliance on a single device for processing.
    In this contribution, we extend a distributed adaptive algorithm for blind system identification that relies on the knowledge of a stacked network-wide solution vector at each node, the computation of which requires either broadcasting or relaying of node-specific values to all other nodes.
    The extended algorithm employs a distributed-averaging-based estimation scheme to estimate the network-wide norm value by only using the information provided by neighboring sensor nodes.
    We introduce an adaptive mixing factor between instantaneous and recursion values for adaptivity in a time-varying system.
    Simulation results show that the extension provides estimation results close to the optimal fully-connected-network case while reducing inter-node transmission significantly.
\end{abstract}
%
\begin{keywords}
multi-channel signal processing, distributed signal processing, wireless sensor networks, blind system identification, distributed averaging
\end{keywords}
%
\section{INTRODUCTION}
\label{sec:intro}

Distributed algorithms have been an active area of research for quite some time, with numerous control, optimization, and signal processing applications.
With the ever-growing number of smart multimedia devices in today's surroundings providing ubiquitous processing and communication capabilities, distributed audio and speech signal processing have also found their way into the spotlight.
Algorithms for distributed signal estimation \cite{5483092}, noise control and echo cancellation \cite{9670697}, as well as beamforming \cite{6663655,6329934,MARKOVICHGOLAN20154} to name a few, have been proposed.
The task of distributed single-input-multiple-output (SIMO) blind system identification (BSI) had contributions such as the adaptive cross-relation-based (CR) \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016}.
Furthermore, we recently introduced an adaptive CR-based algorithm \cite{blochbergerDBSI} using the alternating direction method of multipliers (ADMM) \cite{boydDistributedOptimizationStatistical2011}.
All of the mentioned distributed BSI algorithms rely on information shared between neighboring sensor nodes within the network.
However, the CR-based BSI task necessitates a non-triviality constraint on the full system to be identified (we refer the reader to, e.g., \cite{huangAdaptiveMultichannelLeast2002,huangClassFrequencydomainAdaptive2003}), which manifests itself as one or more global variables.
In this case, the computation of the global variable requries information from all the nodes within the network.
The algorithm in \cite{blochbergerDBSI} relies on node-wise values being relayed throughout the network.
In contrast, to overcome the need for the network to be fully connected, \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016} use an average consensus \cite{xiaoFastLinearIterations2004} approach where a secondary recursion estimates the global variable for each signal frame.
Both approaches introduce additional transmissions of variables between nodes, the number of which, depending on network and neighborhood size, can be substantial or even unfeasable.

In this paper, we extend \cite{blochbergerDBSI} with a distributed averaging-based \cite{xiaoFastLinearIterations2004} estimation scheme for the global variable.
This approach allows us to compute the global variable without the need of a fully connected network or broadcasting.
Further, we introduce a mixing factor to include instantaneous values into the averaging recursion, which then allows us (i) to reduce the number of secondary iterations significantly (down to 1) and (ii) track time-varying systems.
Simulation results show how the extension leads to BSI performance close to an optimal case where all information is available without the need of broadcasting variables to all nodes or a large number of estimation iterations.

\section{DISTRIBUTED ADAPTIVE BSI IN SENSOR NETWORKS WITH ONLINE-ADMM}
\label{sec:dbsi}
In this section, we briefly outline the adaptive SIMO BSI algorithm proposed in [8], setting the scene for the distributed averaging scheme proposed herein.
% In this section, we will introduce certain parts of an adaptive SIMO BSI algorithm based on Online-ADMM \cite{blochbergerDBSI}, which are necessary for the reader to understand its extension of it.
We refer the reader to the publication cited above for the full derivation of the algorithm.

\subsection[]{Sensor network}
We assume a set \(\Mset \triangleq \{1,\ldots,M\}\) sensor nodes indices and a set of edge tuples \(\mathcal{E}\), which connect the nodes forming a sensor network.
Each edge is an unordered pair of node indices \(\{i,j\} \in \mathcal{E}\), which represents the communication link between them.
We allow \(\{i,i\} \in \mathcal{E}\), in order to simplify notation later, however, this does not represent a link but rather indicates the obvious fact that node \(i\) has information of itself.
Furthermore, let \(\Nset_i = \{j|\{i,j\} \in \mathcal{E}\}\) denote the neighborhood of node \(i\),
% From this follows the neighborhood of a node \(i\), i.e., the other nodes it is directly connected to, defined as \(\Nset_i = \{j|\{i,j\} \in \mathcal{E}\}\).
% It is to note, that we define the set with \(i \in \Nset_i\) for ease of notation later.
% This does however not represent an actual inter-node link.
We can define the symmetric adjacency matrix \(\mtxb{C}\) with elements
\begin{equation}
    C_{ij} = \begin{cases}
        1 \quad \text{if } \{i,j\} \in \mathcal{E}\\
        0 \quad \text{otherwise}.
    \end{cases}
\end{equation}
It may be noted that in \cite{blochbergerDBSI}, the pairs \(\{i,j\} \in \mathcal{E}\) are ordered, which yields a directed graph and leads to two sets of neighborhoods ("transmit" and "receive") and therefore a non-symmetric adjacency matrix.
For the sake of brevity, we limit the explanations in this paper to the undirected case.
The case of directed graphs can be extended directly from the presented results.

\subsection[]{Signal model}
We consider a SIMO system with 
\begin{align}
    \mtxb{s}(n) &= [s(n),\,\ldots,\,s(n-2L+2)]^{\T},\\
    \x_i(n) &= [x_i(n),\,\ldots,\,x_i(n-L+1)]^{\T}, \quad i \in \Mset,
\end{align}
the input signal frame and \(M\) output signal frames, respectively.
Each output \(\x_i(n)\) is the convolution of \(\mtxb{s}(n)\) with the respective channel impulse response \(\h_i\) and an additive noise term \(\mtxb{v}_i(n)\), assumed to be zero-mean and uncorrelated with \(\mtxb{s}(n)\).
The signal model is \(\x_i(n) = \mtxb{H}_i \mtxb{s}(n) + \mtxb{v}_i(n),\)
with \(\mtxb{H}_i\), the \(L \times (2L-1)\) linear convolution matrix of the \(i\)th channel using the elements of \(\h_i\) of length \(L\).
For the purpose of this paper, the length \(L\) of the impulse responses is assumed to be known.

\subsection[]{Distributed BSI with Online-ADMM}
In BSI, the coss-relation problem formulation only uses output signals \(\x_i(n)\), exploiting relative information between them, to identify the system, i.e., the acoustic or communication channels \(\h = [\h_1^\T,\,...,\,\h_M^\T]^\T\).
The solution to this problem is found by the minimization problem (cf. e.g. \cite{langtongBlindIdentificationEqualization1994,huangAdaptiveMultichannelLeast2002,huangClassFrequencydomainAdaptive2003,blochbergerDBSI})
\begin{equation}
    \begin{aligned}
        \arg \min_{\h} \quad &\h^\herm \R \h \\
        \text{s.t. } \quad &\|\h\| = 1.
    \end{aligned}\label{eq:frequency_domain:min_prob}
\end{equation}
% Separating the cost function and applying general-form consensus ADMM as an adaptive algorithm leads to the update steps
% \begin{align}
%     \w_i^{n+1} &= \underset{\w_i}{\operatorname{argmin}} \, \mathcal{L}_{\rho} (\w,\h^n,\uu^n)\label{eq:general_consensus_admm:local}\\
%     \h^{n+1} &= \underset{\h, \|\h\| = 1}{\operatorname{argmin}}\, \mathcal{L}_{\rho} (\w^{n+1},\h,\uu^n)\label{eq:general_consensus_admm:global}\\
%     \uu_i^{n+1} &= \uu_i^{n} + \rho \left( \w_i^{n+1} - \tilde{\h}_i^{n+1} \right),\label{eq:general_consensus_admm:dual}
% \end{align}
% where \(\w_i\) represents channels estimates of lower dimensional subproblems, solved locally by node \(i\), \(\h = [\hat{\h}_1^\T,\,...,\,\hat{\h}_M^\T]^\T\) the consensus variable (estimate of the system), and \(\uu_i\) are the dual variables.
\begin{todo}
    find a good way to introduce equation of interest without having to explain all the local, dual variables and parameter mapping in detail.
    \(\h\) notation is confusing.
\end{todo}
The updates of \(\hat{\h}_i\) are of interest for this extension.
It is defined \cite{blochbergerDBSI} as
\begin{equation}
    \hat{\h_i}^{n+1} = \frac{\bar{\h}_i^{n+1}}{\sqrt{\sum_{j \in \Mset} \|\bar{\h}_j^{n+1}\|^2}}.\label{eq:online_admm:consensus_update}
\end{equation}
% for each node \(i \in \Mset\) with the local unnormalized consensus \(\bar{\h}_i^{n+1}\), a combination of neigborhood averages of \(\w_j\) and \(\uu_j\), \(j \in \Nset_i\).
The computation of the denominator of \eqref{eq:online_admm:consensus_update} requires  \(\|\bar{\h}_j^{n+1}\|\) from all nodes \(j \in \Mset\), which however, is not possible without a fully-connected network or broadcasting.
In \cite{blochbergerDBSI}, we assume that partial squared norms are relayed throughout the network until all nodes have the information necessary.
In \cite{yuDistributedBlindSystem2014,liuDistributedBlindIdentification2016}, a global value is estimated by a distributed averaging approach, iteratively combining values within neighborhoods until convergence to a network-wide average.

Subsequently, we introduce an extension to the algorithm described in \cite{blochbergerDBSI} using a fastest distributed linear averaging (FDLA) approach \cite{xiaoFastLinearIterations2004} to avoid the need of network wide data transmission, with the further introduction of an adaptive mixing factor to include instantaneous values in the recursion.

\section{Adaptive norm estimation}
\label{sec:adaptivenormest}

\subsection[]{Distributed averaging}
The computation of the denominator of \eqref{eq:online_admm:consensus_update} requires knowledge of all \(\|\bar{\h}_i\|^2\) at each node \(i \in \Mset\).
To avoid the large number of transmission, which would be necessary to facilitate this, we will approximate the squared norm \(\sum_{i \in \Mset} \|\bar{\h}_i\|^2\) at each node \(i\) only using information shared within its neighborhood \(\Nset_i\).

We look at distributed linear combinations with the iteration index \(k=0,...,K\),
\begin{equation}
    \phi_i({k+1}) = \sum_{j \in \Nset_i} W_{ij} \phi_j({k}),\quad i\in \Mset,\label{eq:adaptivenormest:distlincomb}
\end{equation}
which can be written in vector form as \(\bm{\phi}({k+1}) = \W \bm{\phi}({k})\) with \(\bm{\phi}(k) = \begin{bmatrix} \phi_1(k) & \ldots & \phi_M(k) \end{bmatrix}^\T\).
We want to find the matrix \(\mtxb{W}\), which ensure that for any initial value \(\bm{\phi}({0})\), each element of \(\bm{\phi}({k})\) converges to the average of the elements of \(\bm{\phi}({0})\) when \(k \to \infty\).
In this application, the elements of \(\bm{\phi}(0)\) are the initial node-wise squared norm values \(\|\bar{\h}_i\|^2\) which then leads to \(\lim_{k \to \infty} M \phi_i(k) = \sum_{i \in \Mset} \|\bar{\h}_i\|^2\) for all \(i \in \Mset\).
This is achieved by solving the fastest distributed linear averaging (FDLA) problem, introduced in \cite{xiaoFastLinearIterations2004}, in the form of
\begin{equation}
    \begin{aligned}
        \min& \quad \| \W - \bm{1}\bm{1}^\T/M\|_2\\
        \text{s.t.}& \quad \W \in \mathcal{C},\, \bm{1}^\T \W = \bm{1}^\T,\, \W \bm{1}= \bm{1}.
    \end{aligned}\label{eq:adaptivenormest:fdlaminprob}
\end{equation}
Here, the set \(\mathcal{C} = \{\W \in \mathbb{R}^{M \times M} \vert W_{ij} = 0 \text{ if } \{i,j\} \notin \mathcal{E}\}\) describes all matrices with the same sparsity pattern as the adjacency matrix \(\mtxb{C}\).
Now, the \(i\)-th row of \(\W\) contains the neigborhood weights for node \(i\) as non-zero entries.
% \begin{todo}
%     This section reads awkward
% \end{todo}
% In usual applications of the iteration \eqref{eq:adaptivenormest:distlincomb}, it is applied \(K \in \mathbb{N}\) times until convergence of the averages is reached.
In an adaptive algorithm, where the resulting variable is required at each time frame \(m\) such as e.g. \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016} and \cite{blochbergerDBSI}, it introduces \(K\) secondary iterations per frame.
Each iteration requires information exchange within neighborhoods, adding to the communication cost of the algorithm.
In order to mitigate this, we split iterations over frames \(n\) to keep \(K\) as low as possible.
However, as each frame introduces new data, we need to include this data into the recursion.
To this end, we further introduce an instantaneous weighted neighborhood average of node-wise squared norm values \(\bar{\phi}_i^{n} = \sum_{j \in \Nset_i} W_{ij} \|\bar{\h}_i^n\|^2\) which is then combined with the iterative average into
\begin{equation}
    \hat{\phi}_i^{n+1} = \gamma_i \bar{\phi}_i^{n} + (1-\gamma_i) \phi_i^{n}(K)
\end{equation}
where \(\gamma_i\) is a mixing factor.
With the right choice of mixing factor \(\gamma_i\), this leads to a reduction of distributed averaging iterations \(K\) (see \autoref{sec:simulations}) and improves behaviour of the algorithm when the norms \(\|\h_i\|\) are time-variant.
% The factor balances instantaneous values and the distributed averaging iterations, and we shoudl look at the two extreme cases for further insight.
% For \(\gamma_i = 0\), only instantaneous neighborhood averages are used, which effectively leads to suboptimal computation, as \(\hat{\phi}_1^{n} = \hat{\phi}_2^{n} = \ldots = \hat{\phi}_M^{n}\) for large enough \(m\) will never be reached.
% For \(\gamma_i=1\), only the distributed averaging iterations are considered, i.e., we effectively combine the per-frame secondary recursion \eqref{eq:adaptivenormest:distlincomb} from \(K\) to \(mK\) iterations.
% There however, no new data is introduced which leads to performance degradadation in the adaptive algorithm as it only considers inital estimates from \(m=1\), which might be far of the optimum.

\subsection[]{Adaptive mixing factor}
To avoid choosing a fixed value for \(\gamma_i\), we use information provided by \(\bar{\phi}_i^{n}\).
If the absolute difference between subsequent frames \(| \bar{\phi}_i^{n} - \bar{\phi}_i^{n-1} |\) nears 0, then this means that the estimation of \(\bar{\h}_i\) is converging to a steady state.
Is this the case, then the emphasis of the algorithm should lie on norm estimation, i.e. the distributed averaging recursion, which in turn means \(\gamma_i\) should go towards 0 as well.
We set \(\gamma_i^{n}\) proportional to the absolute difference between subsequent frames and set an upper limit at 1,
\begin{equation}
    \gamma_i^{n} = \min \left\lbrace \frac{| \bar{\phi}_i^{n} - \bar{\phi}_i^{n-1} |}{\bar{\phi}_i^{n-1}},\,1\right\rbrace.\label{eq:adaptivenormest:adaptivegamma}
\end{equation}
The combination of instantaneous values and distributed averaging estimate should lead to lower steady-state error when \(\gamma_i\) is chosen right, while with adaptive \(\gamma_i\), convergence speed should increase as well.
\autoref{alg:davg_norm_est} summarizes the procedure as is is introduced in this section.

\begin{algorithm}[t]
    \caption{ADMM BSI with FDLA-based adaptive estimation of norm values}\label{alg:davg_norm_est}
    \(\W \gets\) \eqref{eq:adaptivenormest:fdlaminprob}\;
    \(\bar{\phi}_i^{0} \gets 1, \forall i \in \Mset\)\;
    \For(){\(n=1\dots\)}
    {
        \For(){\(i \in \Mset\)}
        {
            % \(\cdots\)\\
            % \(\w_i^n \gets \operatorname{argmin}_{\w_i}\, \mathcal{L}_{\rho} (\w,\h^n,\uu^n)\)\;
            % Compute \(\bar{\h}_i^{n}\)\;
            % \(\cdots\)\\
            \emph{The steps before are as introduced in }\cite{blochbergerDBSI}\\
            \dotfill\\
            Transmit \(\bar{\h}_i^{n}\) to nodes  \(j \in \Nset_i\)\;
            \(\bar{\phi}_i^{n} \gets \sum_{j \in \Nset_i} W_{ij} \|\bar{\h}_i^n\|^2\)\;
            \(\gamma_i^{n} \gets \min \left\lbrace \frac{| \bar{\phi}_i^{n} - \bar{\phi}_i^{n-1} |}{\bar{\phi}_i^{n-1}},\,1\right\rbrace\)\;
            \eIf(){\(m = 1\)}
            {
                \(\phi_i^{n}(1) \gets \|\bar{\h}_i^n\|^2\)\;
            }
            {
                \(\phi_i^{n}(1) \gets \hat{\phi}_i^{n-1}(K)\)\;
            }
            \For(){\(k=1,\dots,K\)}
            {
                Transmit \(\phi_i(k)\) to nodes  \(j \in \Nset_i\)\;
                \(\phi_i({k+1}) \gets \sum_{j \in \Nset_i} W_{ij} \phi_j({k})\)\;
            }
            \(\hat{\phi}_i^{n} \gets \gamma_i \bar{\phi}_i^{n} + (1-\gamma) \phi_i^{n}(K)\)\;
            \(\hat{\h_i}^{n} \gets \frac{\bar{\h}_i^{n}}{\sqrt{\hat{\phi}_i^{n}M}}\)\;
            \dotfill\\
            \emph{The steps after are as introduced in }\cite{blochbergerDBSI}\\
            % \(\cdots\)\\
            % \(\uu_i^{n} \gets \uu_i^{n-1} + \rho \left( \w_i^{n} - \tilde{\h}_i^{n} \right)\)\;
            % \(\cdots\)\\
            % % \,\\
        }
    }
\end{algorithm}

% \section{Transmission cost}

\section{Evaluation}
\label{sec:simulations}
\subsection[]{Communication cost}
\label{sec:transcost}
% \begin{table}[t]
%     \centering
%     \begin{tabular}{ |l|l|l| } 
%         \hline
%         & Transmit & Receive \\
%         \hline\hline
%         (1) Full & \(\mathcal{O}(M-1)\) & \(\mathcal{O}(M-1)\) \\
%         \hline
%         (2) Broadcast & \(\mathcal{O}(1)\) & \(\mathcal{O}(M-1)\) \\ 
%         \hline
%         (3) Neigborhood & \(\mathcal{O}(\bar{N}_i K)\) & \(\mathcal{O}(\bar{N}_i K)\) \\ 
%         \hline
%     \end{tabular}
%     \caption[]{Comparison of communication cost.}
%     \label{tab:transcost:table}
% \end{table}
\renewcommand{\arraystretch}{1.2}
\begin{table}[t]
    \centering
    \begin{tabular}{ |l|l|l| }
        \hline
        & Transmit & Receive \\
        \hline\hline
        (1) Fully connected & \(\mathcal{O}(M^2)\) & \(\mathcal{O}(M^2)\) \\
        \hline
        (2) Broadcast & \(\mathcal{O}(M)\) & \(\mathcal{O}(M^2)\) \\ 
        \hline
        (3) Neigborhood & \(\mathcal{O}(M \bar{N}_i K)\) & \(\mathcal{O}(M \bar{N}_i K)\) \\ 
        \hline
    \end{tabular}
    \caption[]{Comparison of communication cost.}
    \label{tab:transcost:table}
\end{table}
\renewcommand{\arraystretch}{1.0}
To describe the communication cost within the network, we adopt the Big-\(\mathcal{O}\) notation, similar to the complexity analysis of algorithms.
We consider the communication of a variable from one node to another as \(\mathcal{O}(1)\).
The most useful measure in the case of this work is to analyze the cost per node.
For this we define an average neighborhood size \(\bar{N}_i = \frac{1}{M} \sum_{i \in \Mset} N_i\) where \(N_i = | \Nset_i |\) is the cardinality/size of the neigborhood set for node \(i\).
We compare the following communication schemes:
\begin{itemize}
    \itemsep-0.2em
    \item[(1)] A fully connected network, i.e., all nodes are linked with all other nodes by direct links, \(\Nset_i = \Mset\).
    \item[(2)] The nodes employ broadcasting, i.e., the nodes transmit to all other nodes without direct links.
    \item[(3)] The nodes only communicate with neighboring nodes, \(\Nset_i \subset \Mset\).
\end{itemize}
For (1), the cost increases quadratically with increasing network size \(M\), both in transmit and receive operations, whereas for (2), the quadratic increase happens only at the latter.
Using communication scheme (3), as in \cite{yuDistributedBlindSystem2014,liuDistributedBlindIdentification2016} and proposed in this paper, cost in proportional to neighborhood size and iteration count \(K\).
Refer to \autoref{tab:transcost:table} for the collected orders of communication cost.

% \begin{figure}[t]
%     \centering
%     \input{simulations/plots/transcostnode/transcostnode.pgf}
%     % \vspace*{-0.8cm}
%     \caption[]{A graphical representation of the order of transmission cost of a single node within the network dependent on network size. One can observe that for the neigborhood-only communication scheme, cost stays constant. (This plot might be unnecessary and space consuming)}
%     \label{fig:transcost:bigo}
% \end{figure}

\subsection[]{Simulation}
\begin{figure}[t]
    \centering
    \input{simulations/plots/icassp2023-static-time/icassp2023-static-time.pgf}
    \caption[]{Median NPM of 30 randomized runs over time for optimal case where all node information is available (fully connected network/broadcasting), fixed values of \(\gamma_i\) with \(K \in \{1,10\}\), adaptive \(\gamma_i\) with \(K=1\). The adaptive method yields results very close to the optimal method. A moving average filter was applied to result curves for better readability (compare transparent and opaque lines). Hatched area indicates frames used in computations for \autoref{fig:simulations:avgNPMgamma}.}
    \label{fig:simulations:NPMtime}
\end{figure}
\begin{figure}[t]
    \centering
    \input{simulations/plots/icassp2023-static/icassp2023-static.pgf}
    \caption[]{Mean over time of median NPM of 30 randomized runs after convergence for varied values for fixed \(\gamma_i\) on the interval \([0.0, 0.04]\) and \(K \in \{1,2,10\}\). Results for optimal method, and adaptive \(\gamma_i\) included for comparison. Hatched area in \autoref{fig:simulations:NPMtime} indicates frames used for computation of avg. NPM.}
    \label{fig:simulations:avgNPMgamma}
\end{figure}
To evaluate the effectiveness of the proposed method, simulations were run.
For this we define the error measure, the normalized projection misalignment
\begin{equation}
    \begin{aligned}
        \text{NPM}(m) &= 20\,\log_{10} \frac{\left\| \mtxb{e} \right\|}{\left\|\h(m)\right\|},\\
        \text{with}\quad \mtxb{e} &= \h(m) - \frac{\h^\T (m) \h_{\text{t}}}{\h_{\text{t}}^\T \h_{\text{t}}}\h(m),
    \end{aligned}
\end{equation}
commonly used in BSI to compare estimated \(\h\) and true \(\h_t\).
The simulation setup is a network with \(M=5\) nodes arranged in a ring topology, each node having \(|N_i|=2\) neighbors.
The input signal is zero-mean white Gaussian noise (WGN), the impulse responses of length \(L=16\) are drawn from a normal distribution, and at each channel, independent WGN is added with \(\text{SNR}=10\,\text{dB}\).
The norms of the impulse responses are scaled to random values drawn from the uniform distribution \(\mathcal{U}_{[0.5,2.0]}\).
A comparison is made between the following cases:
\begin{itemize}
    \itemsep-0.2em
    \item[(a)] all information is available for global norm computation ("optimal" in \autoref{fig:simulations:NPMtime} \& \autoref{fig:simulations:avgNPMgamma}),
    \item[(b)] neighborhood communication with fixed \(\gamma_i \in [0.0, 0.4]\) and \(K \in \{1,2,10\}\),
    \item[(c)] neighborhood communication with adaptive \(\gamma_i\) \eqref{eq:adaptivenormest:adaptivegamma} and \(K=1\).
\end{itemize}
\autoref{fig:simulations:NPMtime} shows the median of 30 Monte-Carlo runs.
We can observe that when fixing \(\gamma_i\), a tradeoff between convergence speed and steady-state error arises at low iteration counts (cf. also \autoref{fig:simulations:avgNPMgamma}).
With \(K=10\), the performance comes close to the optimal case at the cost of additional communication between nodes.
Compare that to the case with adaptive \(\gamma_i\) where even at \(K=1\), i.e., no additional communication for the recursive estimation scheme, convergence speed and steady-state error are close to the optimal case.

\begin{todo}
    TODO: dynamic scenario: add comparisons with optimal and static \(\gamma\) case, and describe in text, add figure captions, extend conclusion
\end{todo}

% \begin{figure}[t]
%     \centering
%     \input{simulations/plots/icassp2023-dynamic-time/icassp2023-dynamic-time.pgf}
%     \caption[]{}
%     \label{fig:simulations:NPMtimedyn}
% \end{figure}
% \begin{figure}[t]
%     \centering
%     \input{simulations/plots/icassp2023-dynamic-norm/icassp2023-dynamic-norm.pgf}
%     \caption[]{}
%     \label{fig:simulations:normtime}
% \end{figure}
% \begin{figure}[t]
%     \centering
%     \input{simulations/plots/icassp2023-dynamic-norms/icassp2023-dynamic-norms.pgf}
%     \caption[]{}
%     \label{fig:simulations:normstime}
% \end{figure}

\begin{figure}[t]
    \centering
    \input{simulations/plots/icassp2023-dynamic-norm/icassp2023-dynamic-norm.pgf}\\\vspace*{-1.0cm}
    \input{simulations/plots/icassp2023-dynamic-norms/icassp2023-dynamic-norms.pgf}\\\vspace*{-1.0cm}
    \input{simulations/plots/icassp2023-dynamic-time/icassp2023-dynamic-time.pgf}
    \caption[]{}
    \label{fig:simulations:NPMtimedyn}
\end{figure}

\section[]{Conclusions}
\label{sec:conclusions}
In this contribution, we propose an extension to a distributed adaptive BSI algorithm applied in sensor networks.
Based on distributed averaging, only using the information provided by neighboring nodes in the sensor network, we compute estimates of norm values of channel estimates in order to enforce a norm constraint.
By balancing the averaging with introducing new data into the recursion, we allow the algorithm to follow an adaptive updating scheme.
The mixing factor, which balances new data with recursion data, is set adaptively dependent on instantaneous channel estimate norms.

These introductions reduce inter-node transmissions for each time frame and improve steady-state estimation performance to values close to the optimal case, where network-wide information is available at all nodes.

% Below is an example of how to insert images. Delete the ``\vspace'' line,
% uncomment the preceding line ``\centerline...'' and replace ``imageX.ps''
% with a suitable PostScript file name.
% -------------------------------------------------------------------------
% \begin{figure}[htb]

% \begin{minipage}[b]{1.0\linewidth}
%   \centering
%   \centerline{\includegraphics[width=8.5cm]{image1}}
% %  \vspace{2.0cm}
%   \centerline{(a) Result 1}\medskip
% \end{minipage}
% %
% \begin{minipage}[b]{.48\linewidth}
%   \centering
%   \centerline{\includegraphics[width=4.0cm]{image3}}
% %  \vspace{1.5cm}
%   \centerline{(b) Results 3}\medskip
% \end{minipage}
% \hill
% \begin{minipage}[b]{0.48\linewidth}
%   \centering
%   \centerline{\includegraphics[width=4.0cm]{image4}}
% %  \vspace{1.5cm}
%   \centerline{(c) Result 4}\medskip
% \end{minipage}
% %
% \caption{Example of placing a figure with experimental results.}
% \label{fig:res}
% %
% \end{figure}

\vfill\pagebreak

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\bibliographystyle{IEEEbib}
\bibliography{refs}

\end{document}